---
title: "Assignment 1: Text Pre-Processing & Clustering"
author: "Stefan Haböck & Bernhard Mayr"
date: "15.11.2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dependencies
```{r dependencies, echo = T, results = 'hide', warning=FALSE}
library(tm)
library(cluster)
library(factoextra)
library(proxy)
library(dplyr)
library(purrr)
```

## Loading Data
- `strip.white=TRUE` automatically trims the document contents
- `tibble` is the newer DataFrame alternative
- `select(...)` prepares the data for the `tm`-package

```{r data loading}
loaded_lectures <- read.table("./lectures.txt", sep="\t", header=TRUE, strip.white=TRUE) %>%
  tibble::as_tibble() %>%
  select(doc_id = ID, text = Description, title = Title) %>%
  DataframeSource() %>%
  VCorpus()
```

## Data Transformation
- first transform the words to lowercase for more equality and because our clustering approach does not differentiate between uppercase and lowercase words
- then apply stopword removal
- and then stem the words

```{r transformations}
prepared_lectures <- loaded_lectures %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removeWords, stopwords::stopwords("english")) %>%
  tm_map(stemDocument)
```

## 

```{r matrices}
dtm <- DocumentTermMatrix(prepared_lectures)
dtm.tfidf <- dtm %>%
  weightTfIdf() %>%
  removeSparseTerms(0.99)

tfidf.matrix <- as.matrix(dtm.tfidf)
dist.matrix = dist(tfidf.matrix, method = "cosine")
```

```{r clustering}
truth.K = 8
clustering.kmeans <- kmeans(tfidf.matrix, truth.K)
clustering.hierarchical <- hclust(dist.matrix, method = "ward.D2")

master.cluster <- clustering.kmeans$cluster 
slave.hierarchical <- cutree(clustering.hierarchical, k = truth.K)
```

```{r plotting, warning=FALSE}
# Silhouette Plot
kmm <- kmeans(tfidf.matrix, truth.K)
D <- daisy(tfidf.matrix, metric = "gower")
plot(silhouette(kmm$cluster, D), col=1:truth.K, border=NA)

# Cluster Dendrogram
plot(clustering.hierarchical, hang = -1)
rect.hclust(clustering.hierarchical, k = truth.K, border = "red")
```

```{r clustered plotting}
points <- cmdscale(dist.matrix, k = 2) 
palette <- colorspace::diverge_hcl(truth.K) # Creating a color palette 
previous.par <- par(mfrow=c(2,2), mar = rep(1.5, 4)) 
 
plot(points, main = 'K-Means clustering', col = as.factor(master.cluster), 
     mai = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), 
     xaxt = 'n', yaxt = 'n', xlab = '', ylab = '') 
plot(points, main = 'Hierarchical clustering', col = as.factor(slave.hierarchical), 
     mai = c(0, 0, 0, 0), mar = c(0, 0, 0, 0),  
     xaxt = 'n', yaxt = 'n', xlab = '', ylab = '')

par(previous.par) # recovering the original plot space parameters
```

## Exercise Answers
- **3c:** Use different values for the number of clusters k. Have a look at the documents
combined to a cluster. Which k works best?
- **3d:** Use different distance measurements combined with different values for k. Which
distance works best (with which cluster size)?
- **3e:** Perform various runs. Does the result look similar? Does changing the number of
iterations have any effect?
- **4e:** Test various distance measurements and different linkage structures. Which work
best? Can you see any peculiarities of the linkage methods?
- **5a:** Use tfidf instead of term occurrence for the TtD. Does it improve the results?

## Resources
- https://mran.microsoft.com/snapshot/2018-03-30/web/packages/tm/vignettes/tm.pdf
- https://cran.r-project.org/web/packages/tm/tm.pdf
- https://medium.com/@SAPCAI/text-clustering-with-r-an-introduction-for-data-scientists-c406e7454e76
- https://books.psychstat.org/textmining/cluster-analysis.html